---
title: "Loan Application Analysis"
author: "Arie Twigt"
output:
  word_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---


```{r, echo=FALSE}
data <- read.csv("~/Projects/r_training/data/german_data_clean.csv")
```


# Data analysis


## Dimensions


The dimensions of the dataset

```{r}
dim(data)
nrow(data)
ncol(data)
```


The data set has the following columns:

```{r}
colnames(data)
length(colnames(data))
```

## Statistics of the dataset

### Entire data set


A summarization of the entire data set


```{r}
summary(data)
```


### Specific columns


Purposes for loans

```{r}
summary(factor(data$purpose))
```


```{r}
barplot(summary(factor(data$purpose)))
```



Mean duration of loans

```{r}
mean_months <- mean(data$duration_months)
mean_months

```


Total number of loans with a duration higher than 20

```{r}
sum(data$duration_months > 20)
```



# Sub-selections of DataFrames


Sub-select a specific cell:

```{r}
data[1,4]
```


Select multiple rows, and columns by name.


```{r}
data[5:20, c("duration_months" ,"purpose", "response")]
```


# Filtering Data Frames

## Individual filters

Loans with a duration higher than 50 months

```{r}

data_high_duration <- data[data$duration_months > 50, ]
data_high_duration
```


What are the purposes of these high duration loans?


```{r}
summary(factor(data_high_duration$purpose))
```


## Combined filters


```{r}

data_high_amount_radio <- data[(data$credit_amount > 5000) & (data$purpose == "radio/television")  ,]
data_high_amount_radio
```


## Creating categories in R


```{r}

data$age_class <- cut(data$age_years, 
                      breaks = c(18, 25, 50, 65, 100), 
                      labels = c("young", "adult", "senior", "retired")) 
```

Only selecting the age_years and age_class columns

```{r}
data[,c("age_years", "age_class")]
```


Display how much values we have for each category:


```{r}
table(data$age_class)
```


# Exploration with basic visualisations


```{r}
count_age_class <- table(data$age_class)

barplot(count_age_class, main = "Loans by age class",
                         xlab = "Age class",
                         ylab = "Count",
                         col = "#fa6400",
                         space = 0)
```


Boxplot


```{r}
boxplot(data$credit_amount,
        col = "#fa6400",
        outline=FALSE,
        notch = TRUE)
```


Piecharts


```{r}
pie(count_age_class, 
    main="Age category",
    sub="By our own categories",
    col = c("green", "blue", "yellow", "black"))
```



Histogram


```{r}
hist(data$credit_amount, breaks=500)
```


```{r}
plot(density(data$age_years))
```

## Ordering the dataset


```{r}
data[order(data$duration_months, decreasing = TRUE), ]
```

